The weighted hyperdirichlet distribution: the `hyper3` class

### 2025-02-08 inst/hyper3.Rmd
expectations

```{r setup, include=FALSE}
set.seed(1)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
```

@@

$$
\hat{p_1}=\frac{\sqrt{ac}}{\sqrt{ac}+\sqrt{bd}}\qquad
\hat{p_2}=\frac{\sqrt{bd}}{\sqrt{ac}+\sqrt{bd}}\qquad
\hat{\lambda}=\sqrt{\frac{ad}{bc}}
$$
Likelihoods of this form are easily specified using package idiom.

>>>>

```{r setup, include=FALSE}
set.seed(1)
knitr::opts_chunk$set(echo = TRUE)
library("hyper2")
library("magrittr")
```

@@

$$
\hat{p_1}=\frac{\sqrt{ac}}{\sqrt{ac}+\sqrt{bd}}\qquad
\hat{p_2}=\frac{\sqrt{bd}}{\sqrt{ac}+\sqrt{bd}}\qquad
\hat{\lambda}=\sqrt{\frac{ad}{bc}}
$$

If we consider $p_1$ at home and $p_2$ away, we have $a+b$ matches.
Using the maximum likelihood estimates
$\hat{p_1},\hat{p_2},\hat{\lambda}$, the expected number of wins by
$p_1$ is

$$
(a+b)\frac{\hat{\lambda}\hat{p_1}}{\hat{\lambda}\hat{p_1} + \hat{p_2}}
=
(a+b)\cdot
\frac{
\sqrt{\frac{ad}{bc}}\frac{\sqrt{ac}}{\sqrt{ac}+\sqrt{bd}}
  }{
\sqrt{\frac{ad}{bc}}\frac{\sqrt{ac}}{\sqrt{ac}+\sqrt{bd}} + \frac{\sqrt{bd}}{\sqrt{ac}+\sqrt{bd}}
}
=\cdots=a
$$

Also the expected number of matches won by $p_2$ when $p_1$ is at home
can be shown to be $b$, and similarly when $p_2$ is at home, the
expected number of victories is $c,d$ respectively.  Thus, at least in
the two-team case, the expectations under the ML estimates are exactly
the observations.

### `hyper2` idiom for likelihood functions {-}

Likelihoods of this form are easily specified using package idiom.
